---
title: "Take Home Exercise 1"
author: "You Ting Quek"
date: "April 15, 2024"
date-modified: "Last-Modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Background
There are two major residential property market in Singapore, namely public and private housing. Public housing aims to meet the basic need of the general public with monthly household income less than or equal to S$14,000. For families with monthly household income more than S$14,000, they need to turn to the private residential market.

# The Task
Assuming the role of a graphical editor of a median company, you are requested to prepare minimum two and maximum three data visualisation to reveal the private residential market and sub-markets of Singapore for the 1st quarter of 2024

## Data Source
To accomplish the task, transaction data of REALIS (2023-2024) will be used.

# Data visualisation 
## Preparing the Data

## Installing required packages
```{r}
pacman::p_load(tidyverse, ggplot2, dplyr)
```

### Data Preparation

```{r}
# Define the paths to the individual CSV files
file1 <- "data/ResidentialTransaction20240308160536.csv"
file2 <- "data/ResidentialTransaction20240308160736.csv"
file3 <- "data/ResidentialTransaction20240308161009.csv"
file4 <- "data/ResidentialTransaction20240308161109.csv"
file5 <- "data/ResidentialTransaction20240414220633.csv"

# Read the individual CSV files
data1 <- read_csv(file1)
data2 <- read_csv(file2)
data3 <- read_csv(file3)
data4 <- read_csv(file4)
data5 <- read_csv(file5)

# Combine the data frames into one
combined_transaction <- bind_rows(data1, data2, data3, data4, data5)

glimpse (combined_transaction)

```
### Duplicate checks


## Price index for private residential properties for 1st Quarter 2024

## Upgraders from HDB for 1st Quarter 2024