---
title: "Hands_on_ex1"
author: "QUEK You Ting"
date: "April 13, 2024"
date-modified: "Last-Modified"
execute: 
  eval: true
echo: true
warning: false
freeze: true
---

##Step 1: Installing the required libraries

```{r}
pacman::p_load(tidyverse)
```


### Importing the required data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

##Step 2: View the data in histogram

```{r}
hist(exam_data$MATHS)

ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```

##Step 3: Essential Grammatical Elements in ggplot2 with Aesthetic Mapping

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

## Visualising data in Histogram

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

## Visualising data in Dotplot

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

## Visualising data in Histogram in bins

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()       
```

### Adding colors

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

## Visualising data with data plotlines - geomdensity()

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

## Visualising data with Boxplot diagrams

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

## Visualising data with violin plot - geomviolin()

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

## Visualising data in scatterplot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()            
```

## Combining boxplot with scatter plot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             size = 0.5)        
```

## Working with Essential Grammatical Elements in ggplot2: stat

### Working with stat summary

```{r}

ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun="mean",         
               colour ="red",        
               size=4)               

```

### Overriding soothing methods

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5)
```

## Working with Essential Grammatical Elements in ggplot2: Facet

### Using Facetwraps

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

### Using Facet Grid

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```
