---
title: "Hands On Exercise 3"
author: "You Ting QUEK"
date: "April 25, 2024"
date-modified: "Last-Modified"
execute: 
  eval: true
echo: true
warning: false
freeze: true
---

# Objectives:

-   Create interactive data visualisation by using functions provided by **ggiraph** and **plotlyr** packages

# Getting Started

## Installing the required libraries

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse)
```

## Importing the Data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

To create an interactive version of ggplot2 geom (i.e. [`geom_dotplot_interactive()`] we first need to create the basic graph. Then, [`girafe()`](https://davidgohel.github.io/ggiraph/reference/girafe.html) will be used to generate an svg object to be displayed on an html page.

```{r}
#| echo: false

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```