---
title: "In-class_Ex02 Demo"
author: "You Ting QUEK"
date: "April 20, 2024"
date-modified: "Last-Modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Setting up the environment

### Installing required packages

```{r}
pacman::p_load(tidyverse, ggplot2, ggdist, ggthemes, colorspace, ggridges)
```

### Importing the Exam_data.csv

```{r}
exam_df <- read_csv("data/Exam_data.csv")
```

## Visualising Distribution

```{r}

median_eng <- median(exam_df$ENGLISH)
mean_eng <- mean(exam_df$ENGLISH)
std_eng <- sd(exam_df$ENGLISH)

ggplot(exam_df,
       aes(x = ENGLISH)) +
  geom_density(
    color = "#1696d2",
    adjust = .65,
    alpha = .6
  )+
  stat_function(
    fun = dnorm,
    args = list(mean = mean_eng,
                sd = std_eng),
    col = "grey30",
    size = .8) +
  geom_vline(
    aes(xintercept = mean_eng),
    
  )
```
## Probability Density Plot

Appropriate method for representing continuous values

```{r}
#| echo: false

ggplot(data=exam_df, aes(x = ENGLISH)) +
  geom_density(color="skyblue", 
               adjust = .65,
               alpha = .1) +
  ggtitle("Probability Density Plot of English scores")
```

## Density Plot with Summary Statistics

The code chunk below includes a probability density plot (violet) and normal distribution plot (gray). The mean (green) and median (blue) of the data is also plotted.


```{r}
median_eng <- median(exam_df$ENGLISH)
mean_eng <- mean(exam_df$ENGLISH)
std_eng <- sd(exam_df$ENGLISH)


# Probability Density Plot
ggplot(data=exam_df, aes(x = ENGLISH)) +
  geom_density(color="violet", 
               adjust = .65,
               alpha = .1) +
  stat_function( # Normal Distribution Plot
    fun = dnorm,
    args = list(mean = mean_eng, sd = std_eng),
    col = "gray",
    linewidth = .8) +
  geom_vline( # Mean line
    aes(xintercept = mean_eng),
    colour = "darkgreen",
    linewidth = .6,
    linetype = "dashed") +
  annotate(geom = "text",
           x = mean_eng -8,
           y = .04,
           label = paste0("Mean: ",
                          round((mean_eng),2)),
           colour="darkgreen") +
  geom_vline( # Median Line
    aes(xintercept = median_eng),
    colour = "navy",
    linewidth = .6,
    linetype = "dashed") +
  annotate(geom = "text",
           x = median_eng +8,
           y = .04,
           label = paste0("Median: ",
                          round((mean_eng),2)),
           colour="navy") +
  ggtitle("Density Plot with Summary Statistics of English Scores")
```

## Ridgeline Plot

[*Ridgeline plot*](https://www.data-to-viz.com/graph/ridgeline.html) (i.e. *Joyplot*) reveals the distribution of a numeric value for several groups. Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale with slightoverlaps

```{r}
#| echo: false

ggplot(exam_df, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English Grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

## Varying Fill Colours along the X-axis

Sometimes we would like to have the area under a ridgeline not filled with a single solid color but rather with colors that vary in some form along the x axis.

```{r}

ggplot(exam_df, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = after_stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [°C]",
                       option = "C") +
  scale_x_continuous(
    name = "English Grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

## Mapping Probabilities directly onto colour

Figure below is plotted by mapping the probabilities calculated by using `stat(ecdf)` which represent the empirical cumulative density function for the distribution of English score.


```{r}
#| echo: false

ggplot(exam_df,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()
```

## Ridgeline Plots with Quantile Lines

Ridgeline plots can be coloured by quantile using [`geom_density_ridges_gradient()`](geom_density_ridges_gradient()), via the calculated `stat(quantile)` 


```{r}
#| echo: false

ggplot(exam_df,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```

Quantiles can also be specified by cut points e.g. 2.5% and 97.5% tails to colour the ridgeline plot.

```{r}
#| echo: false

ggplot(exam_df,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("red", "grey", "navy"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()
```

# Raincloud Plot

A data visualisation technique that produces a half-density to a distribution plot. It got its name due to the likeness of the density plot to a shape of a "raincloud". The raincloud (half-density) plot enhances the traditional box-plot by highlighting multiple modalities (an indicator that groups may exist). The raincloud plot shows where densities are clustered.

It is created by using functions provided by **ggdist** and ggplot2 packages.

## Plotting a Half Eye graph

Plot a Half-Eye graph by using [`stat_halfeye()`](https://mjskay.github.io/ggdist/reference/stat_halfeye.html) of **ggdist** package, producing a Half Eye visualization, which is contains a half-density and a slab-interval.

::: panel-tabset
## Plot

```{r}
#| echo: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA)
```

## Code

```{r}
#| echo: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA)
```
:::

::: callout-note
Remove the slab interval by setting .width = 0 and point_colour = NA.
:::

## **Adding the boxplot**

The second geometry layer i.e. a narrow boxplot is produced using [`geom_boxplot()`](https://r4va.netlify.app/chap09) of ggplot2 This produces a narrow boxplot.

::: panel-tabset
## Plot

```{r}
#| echo: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)
```

## Code

```{r}
#| eval: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)
```
:::

## **Adding the Dot Plots**

The third geometry layer is added using [`stat_dots()`](https://mjskay.github.io/ggdist/reference/stat_dots.html) of ggdist package. This produces a half-dotplot, similar to a histogram that indicates the number of samples (number of dots) in each bin. Use `side = “left”` to specify the dot plots on the left-hand side.

::: panel-tabset
## Plot

```{r}
#| echo: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)
```

## Code

```{r}
#| eval: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)
```
:::

## **Finishing touch**

 [`coord_flip()`](https://ggplot2.tidyverse.org/reference/coord_flip.html) of ggplot2 package is used to flip the raincloud chart horizontally to give it the raincloud appearance. `theme_economist()` of ggthemes package is also used to give the raincloud chart a professional publishing standard look.

::: panel-tabset
## Plot

```{r}
#| echo: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = NA,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```

## Code

```{r}
#| eval: false

ggplot(exam_df, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```
:::
